A Comparative study of Dask and Spark for data-intensive neuroimaging pipelines

### 2019-07-02 latex/paper.tex
Went through Section III

\title{A Comparative study of Dask and Spark for neuroimaging pipelines}

@@

@@

@@

%%% INCREMENTATION %%%
%% worker
\subsection{Incrementation: Number of workers}
Figure~\ref{fig:inc_ms_worker} shows the makespan of the Incrementation
application for different numbers of workers and for different engines.
Overall, there is no substantial makespan difference between the engines.
Dask seems to have a slight advantage over Spark, \TG{X (toutes les Dask vs Spark)} seconds on
average, with Delayed and Futures being slightly better than Bags.

@@

Figure~\ref{fig:inc_ms_block} shows the Incrementation makespan when
varying the number of image blocks for a constant BigBrain image size. We
were not able to run Spark for 30 blocks due to its 2GB limitation in the
task size it can compute. Once again, we do not observe any significant
difference among the engines.

In Figure~\ref{fig:inc_tt_block}, the total execution time of each function is shown.
For 30 blocks the Dask Bag API has a much lower overhead time. This is because we
only calculate the idle time of the used core. Dask Bag was only using one thread per
block in comparison to Dask Delayed and Dask Futures which was offloading a block
calculations on multiple threads of the same worker. Furthermore, the variance of the
overhead increases proportionally to the number of blocks. Also, the IO time reduces
with more blocks however the overhead time increase by a similar amount. This is not
observed for 30 blocks as the workers are not used at full capacity; i.e.\ some
threads are idle.

@@

%% Iterations
\subsection{Incrementation: Number of iterations}
Figure~\ref{fig:inc_ms_itr} shows the makespan of the application for the various
number of iterations. Spark and the Dask Bag API are equivalent in terms of
performance. The Dask Delayed API is slightly faster with few iterations (tasks) but
it has a similar makespan to Spark and Dask Bag when the number of iterations
increases. Overall, Dask Futures seems to be faster but the difference is not
substantial.

In Figure~\ref{fig:inc_tt_itr}, the execution time of the function type is shown. The
compute time increase proportionally to the number of iterations. This is because
each computing task has a constant sleep delay. Also, the overhead increases when the
number of iterations (tasks) increases while the IO time decreases; however this is
not substantial.

@@

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_itr.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_itr}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_itr.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_itr}
    \end{subfigure}
    \caption{125 blocks, 4 sec.\ sleep delay, 8 instances}
\end{figure}

%% Sleep time
\subsection{Incrementation: Sleep delay}
Figure~\ref{fig:inc_ms_sleep} shows the makespan of the incrementation algorithm for
different sleep delays. Spark is initially slower than the Dask APIs, however, it is
faster when increasing the sleep delay. Also, within the Dask APIs, Dask Bag is
slower than the other two but it is not considerable.

Figure~\ref{fig:inc_tt_sleep} shows the execution time for each type of function. On
one hand, Spark has the smallest overhead. Table~\ref{tab:inc_sleep_overhead} shows
that it is substantial; especially with a 64 seconds sleep delay. On the other hand,
Spark has a high IO time. In contrast, Dask Delayed and Futures have the largest
overhead time but the smallest IO time. Then, Dask Bag stands in the middle for both
overhead and IO time.

\begin{table}[!t]
    \caption{Function time for the sleep experiment}
    \begin{subtable}[b]{\columnwidth}
        \renewcommand{\arraystretch}{1.3}
        \caption{Overhead time in second}\label{tab:inc_sleep_overhead}
        \centering
        \begin{tabular}{lllll}
        \hline
                     & 1 sec. & 4 sec. & 16 sec. & 64 sec. \\ \hline
        Spark        & 14992  & 17076  & 16069   & 9769    \\
        Dask.Bag     & 20690  & 21388  & 17759   & 18720   \\
        Dask.Delayed & 27481  & 26199  & 22080   & 24624   \\
        Dask.Futures & 27069  & 27984  & 23538   & 24762   \\ \hline
        \end{tabular}
    \end{subtable}
    \vskip 0.2cm
    \begin{subtable}[b]{\columnwidth}
        \renewcommand{\arraystretch}{1.3}
        \caption{IO time in second}\label{tab:inc_sleep_io}
        \centering
        \begin{tabular}{lllll}
        \hline
                     & 1 sec. & 4 sec. & 16 sec. & 64 sec. \\ \hline
        Spark        & 63622  & 57144  & 51070   & 57078   \\
        Dask.Bag     & 54630  & 52312  & 53743   & 53612   \\
        Dask.Delayed & 45239  & 44142  & 45133   & 46198   \\
        Dask.Futures & 46537  & 43161  & 44749   & 45846   \\ \hline
        \end{tabular}
    \end{subtable}
    \vspace{-3mm}
 \end{table}

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_sleep.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_sleep}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_sleep.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_sleep}
    \end{subfigure}
    \caption{125 blocks, 10 iterations, 8 instances}
\end{figure}

%% Baseline gantt chart
\subsection{Experimentation 1: Baseline timeline}
Figure\ref{fig:inc_gantt} shows the tasks timeline for each worker thread. On the
x-axis is the time in seconds and on the y-axis is the workers and their threads. The
color represents the type of different tasks; note that the color differs from one
graph to another. The first section of reading tasks is much larger than the
following ones. This is because, initially, the NFS is saturated by the number of
workers accessing it. Also, a substantial amount of overhead happens near the IO
tasks. Moreover, for the computing tasks, there is more overhead for the latter tasks.

% The frameworks have similar makespan however the time spend for each type of
% task differs considerably (see Table~\ref{tab:inc_base}). Spark tends to spend more
% time than other frameworks for IO however it has a much lower overhead. Dask Delayed
% and Futures have the lowest IO time but a much higher overhead. Dask Bag stands in
% the middle for both IO and overhead. Compute time is approximately the same for all
% frameworks. Apart from the first IO section, there is a large amount of ovehead
% everytime time IO is performed. This is true for all frameworks.

\begin{figure*}[!htb]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/spark_inc_baseline_gantt.png}
        \caption{Spark execution timeline}\label{fig:inc_spark_gantt}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/Dask_bag_inc_baseline_gantt.png}%
        \caption{Dask Bag execution timeline}\label{fig:inc_dask_bag_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/Dask_delayed_inc_baseline_gantt.png}%
        \caption{Dask Delayed execution timeline}\label{fig:inc_dask_delayed_gantt}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/Dask_futures_inc_baseline_gantt.png}%
        \caption{Dask Futures execution timeline}\label{fig:inc_dask_futures_gantt}
    \end{subfigure}
    \caption{125 blocks, 4 sec.\ sleep delay, 8 instances}\label{fig:inc_gantt}
\end{figure*}

@@

%%% HISTOGRAM %%%
\subsection{Experiment 2: Number of instances}

@@

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_splits.png}%
        \caption{Histogram makespan}\label{fig:histo_ms_block}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_idle_splits.png}%
        \caption{Histogram total time}\label{fig:histo_tt_block}
    \end{subfigure}
    \caption{8 workers}
\end{figure}
\TG{We decided to omit Dask Futures as it does not have any value over Dask
Delayed in this application.}
%%% BIDS EXAMPLE %%%
\subsection{Experiment 3: Number of instances}

@@

>>>>

\title{A Comparative study of Dask and Spark for data-intensive neuroimaging pipelines}

@@

\TG{Say that it is often said that Spark has a larger overhead but scales 
worse than Dask. In the dicussion, get back to it and mention that (1) we don't
always observe that, (2) for data-intensive applications, overheads don't really matter
as reducing them increases data trasnfers.}

@@

For each run, we measured the application makespan as well as the cumulative 
data read, compute, data write, and engine overhead across all application tasks. 
We measured the overhead by \TG{\ldots, maybe add a figure}.

\TG{explain that number of iterations may impact futures.}

@@

%%% INCREMENTATION %%%
%% worker
\subsection{Incrementation: Number of workers}
Figure~\ref{fig:inc_ms_worker} shows the makespan of the Incrementation
application for different numbers of workers and for different engines. 
The bars show the average makespan over 5 repetitions while the error bars 
are the standard deviation \TG{check this}.
Overall, there is no substantial makespan difference between the engines.
Dask seems to have a slight advantage over Spark, \TG{X (toutes les Dask vs Spark)} seconds on
average, with Delayed and Futures being slightly better than Bags.

@@

Figure~\ref{fig:inc_ms_block} shows the Incrementation makespan when
varying the number of image blocks for a constant BigBrain image size. We
were not able to run Spark for 30 blocks due to its 2GB limitation in the
task size it can compute. Once again, we do not observe any significant
difference among the engines. For all engines, makespan variability
increases with the number of blocks, however, engines scale very well in
general.

In Figure~\ref{fig:inc_tt_block}, the total execution time of each function is shown.
For 30 blocks the Dask Bag API has a much lower overhead time. This is because Dask Bag was only using one thread per
block in comparison to Dask Delayed and Dask Futures which were offloading a block
calculations on multiple threads of the same worker.  Once again, the data transfer time reduces
with more blocks however the overhead time increases by a similar amount. This is not
observed for 30 blocks as the workers are not used at full capacity, i.e., some
threads are idle. Finally, the variability of the
overhead increases with the number of blocks, which explains the makespan
variability mentioned previously.

@@

%% Iterations
\subsection{Incrementation: Number of iterations}
Figure~\ref{fig:inc_ms_itr} shows the makespan of the application while
varying the number of iterations. Overall,  Spark and Dask APIs are once
again equivalent, although Dask Delayed and Futures are slightly faster
than Bags and Spark for 1 and 10 few iterations, and Fuures are faster than
Delayed, Bags and Spark for 100 iterations. Differences remain minor
though.

In Figure~\ref{fig:inc_tt_itr}, the total execution time breakdown is
shown. We observe good scalability of all the engines with the number of
iterations.

@@

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_itr.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_itr}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_itr.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_itr}
    \end{subfigure}
    \caption{125 blocks, 4-s delay, 8 workers, 8 threads/worker}
\end{figure}

%% Sleep time
\subsection{Incrementation: Sleep delay}
Figure~\ref{fig:inc_ms_sleep} shows the makespan of the Incrementation
application for different sleep delays. Overall, all engines again perform the same, and scale well with task duration.
 Spark is initially slower than the
Dask APIs, however, it is faster when increasing the sleep delay. Also,
within the Dask APIs, Dask Bag is slower than the other two but it is not
considerable.

Figure~\ref{fig:inc_tt_sleep} shows the total execution time breakdown. On
one hand, Spark has the smallest overhead. As previously observed,
variations in overhead time are almost exactly compensated by variations in data
transfer time.

% \begin{table}[!t]
%     \caption{Function time for the sleep experiment}
%     \begin{subtable}[b]{\columnwidth}
%         \renewcommand{\arraystretch}{1.3}
%         \caption{Overhead time in second}\label{tab:inc_sleep_overhead}
%         \centering
%         \begin{tabular}{lllll}
%         \hline
%                      & 1 sec. & 4 sec. & 16 sec. & 64 sec. \\ \hline
%         Spark        & 14992  & 17076  & 16069   & 9769    \\
%         Dask.Bag     & 20690  & 21388  & 17759   & 18720   \\
%         Dask.Delayed & 27481  & 26199  & 22080   & 24624   \\
%         Dask.Futures & 27069  & 27984  & 23538   & 24762   \\ \hline
%         \end{tabular}
%     \end{subtable}
%     \vskip 0.2cm
%     \begin{subtable}[b]{\columnwidth}
%         \renewcommand{\arraystretch}{1.3}
%         \caption{IO time in second}\label{tab:inc_sleep_io}
%         \centering
%         \begin{tabular}{lllll}
%         \hline
%                      & 1 sec. & 4 sec. & 16 sec. & 64 sec. \\ \hline
%         Spark        & 63622  & 57144  & 51070   & 57078   \\
%         Dask.Bag     & 54630  & 52312  & 53743   & 53612   \\
%         Dask.Delayed & 45239  & 44142  & 45133   & 46198   \\
%         Dask.Futures & 46537  & 43161  & 44749   & 45846   \\ \hline
%         \end{tabular}
%     \end{subtable}
%     \vspace{-3mm}
%  \end{table}

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_sleep.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_sleep}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_sleep.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_sleep}
    \end{subfigure}
    \caption{125 blocks, 10 iterations, 8 workers, 8 threads/worker}
\end{figure}

%% Baseline gantt chart
\subsection{Incrementation: Gantt chart}

Figure~\ref{fig:inc_gantt} shows the Gantt chart obtained for each engine
and API, for a baseline execution with 125 blocks, a 4-second sleep delay
and 8 workers \TG{Make the colors uniform}. Gantt charts are structured in 
batches of up to 64 read-compute-write concurrent sequences.
File reads in the first batch are much longer than in the following ones:
this is due to the high synchronization of data transfers that leads to a
high saturation of the shared file system. We also note that overhead,
represented in white, is concentrated around the data transfer tasks
\TG{find out why}, and the computing tasks that run concurrently with data
transfers. 
% The frameworks have similar makespan however the time spend for each type of
% task differs considerably (see Table~\ref{tab:inc_base}). Spark tends to spend more
% time than other frameworks for IO however it has a much lower overhead. Dask Delayed
% and Futures have the lowest IO time but a much higher overhead. Dask Bag stands in
% the middle for both IO and overhead. Compute time is approximately the same for all
% frameworks. Apart from the first IO section, there is a large amount of ovehead
% everytime time IO is performed. This is true for all frameworks.

\begin{figure*}[!htb]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/spark_inc_baseline_gantt.png}
        \caption{Spark execution timeline}\label{fig:inc_spark_gantt}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/Dask_bag_inc_baseline_gantt.png}%
        \caption{Dask Bag execution timeline}\label{fig:inc_dask_bag_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/Dask_delayed_inc_baseline_gantt.png}%
        \caption{Dask Delayed execution timeline}\label{fig:inc_dask_delayed_gantt}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/Dask_futures_inc_baseline_gantt.png}%
        \caption{Dask Futures execution timeline}\label{fig:inc_dask_futures_gantt}
    \end{subfigure}
    \caption{125 blocks, 4 sec.\ sleep delay, 8 workers}\label{fig:inc_gantt}
\end{figure*}

@@

%%% HISTOGRAM %%%
\subsection{Experiment 2: Number of workers}

@@

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_splits.png}%
        \caption{Histogram makespan}\label{fig:histo_ms_block}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_idle_splits.png}%
        \caption{Histogram total time}\label{fig:histo_tt_block}
    \end{subfigure}
    \caption{8 workers}
\end{figure}
\TG{We decided to omit Dask Futures as it does not have any value over Dask
Delayed in this application.}
%%% BIDS EXAMPLE %%%
\subsection{Experiment 3: Number of workers}

@@

\TG{In the dicussion, get back to it and mention that (1) we don't
always observe that, (2) for data-intensive applications, overheads don't really matter
as reducing them increases data trasnfers.}
